### 2015-03-11 Programming Project C/01knapsack.tex
fixed off by one issue in dynamic solution, added table in paper

%% Based on a TeXnicCenter-Template by Tino Weinkauf.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% HEADER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[onecolumn, 11pt, conference, compsocconf]{IEEEtran}
\documentclass[onecolumn, 12pt, article]{IEEEtran}
\usepackage{times,color,amsmath,amssymb,amsthm,comment,graphicx,cite,multirow}
%\documentclass[letterpaper,oneside,11pt]{letter}
% Alternative Options:
%	Paper Size: a4paper / a5paper / b5paper / letterpaper / legalpaper / executivepaper
% Duplex: oneside / twoside
% Base Font Size: 10pt / 11pt / 12pt

%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[USenglish]{babel} %francais, polish, spanish, ...
\usepackage[T1]{fontenc}
\usepackage[ansinew]{inputenc}

\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}

%% Packages for Graphics & Figures %%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{graphicx} %%For loading graphic files
%\usepackage{subfig} %%Subfigures inside a figure
%\usepackage{tikz} %%Generate vector graphics from within LaTeX

%% Please note:
%% Images can be included using \includegraphics{filename}
%% resp. using the dialog in the Insert menu.
%% 
%% The mode "LaTeX => PDF" allows the following formats:
%%   .jpg  .png  .pdf  .mps
%% 
%% The modes "LaTeX => DVI", "LaTeX => PS" und "LaTeX => PS => PDF"
%% allow the following formats:
%%   .eps  .ps  .bmp  .pict  .pntg

%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{array,MnSymbol}

%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\singlespacing        %% 1-spacing (default)
%\onehalfspacing       %% 1,5-spacing
%\doublespacing        %% 2-spacing

%% Other Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{a4wide} %%Smaller margins = more text per page.
\usepackage{fancyhdr} %%Fancy headings
\usepackage{listings}
\usepackage{capt-of}

%
% Theorem like environments
%
\newtheorem{problem}{Problem}%
%\numberwithin{problem}{section}
\newtheorem{theorem}{Theorem}%
\newtheorem{acknowledgment}{Acknowledgment}%
%\newtheorem{algorithm}{Algorithm}%
\newtheorem{assumption}{Assumption}%
\newtheorem{axiom}{Axiom}%
\newtheorem{case}{Case}%
\numberwithin{case}{problem}
\newtheorem{claim}{Claim}%
\newtheorem{conclusion}{Conclusion}
\newtheorem{condition}{Condition}
\numberwithin{condition}{problem}
\numberwithin{condition}{subsection}
\newtheorem{conjecture}{Conjecture}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}
\newtheorem{definition}{Definition}
\numberwithin{definition}{section}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}%
\newtheorem{lemma}{Lemma}%
\newtheorem{notation}{Notation}%
\theoremstyle{remark}
\newtheorem{question}{Question}%
\numberwithin{question}{problem}
\theoremstyle{plain}
\newtheorem{answer}{Answer}%
\numberwithin{answer}{problem}
\newtheorem{proposition}{Proposition}%
\newtheorem{remark}{Remark}%
\newtheorem{solution}{Solution}%
\numberwithin{solution}{section}
\newtheorem{summary}{Summary}%
\numberwithin{equation}{section}%
\newtheorem{option}{Option}

\raggedbottom
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Remarks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Note:
% 1. Edit the used packages and their options (see above).
% 2. If you want, add a BibTeX-File to the project
%    (e.g., 'literature.bib').
% 3. Happy TeXing!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Options / Modifications
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{options} %You need a file 'options.tex' for this
%% ==> TeXnicCenter supplies some possible option files
%% ==> with its-templates (File | New from Template...).

%% BEGIN DOCUMENT
\begin{document}

%% Title Page
\title{0/1 Knapsack}
\author{Preston Stosur-Bassett}
\date{March 11, 2015}
\maketitle

\pagestyle{fancy}
\fancyhead[R]{0/1 Knapsack in Java, page \thepage}
\fancyhead[L]{Preston Stosur-Bassett}

%% BEGIN ABSTRACT
\begin{abstract}
%%TODO: write this

%% END ABSTRACT
\end{abstract}

%% BEGIN MOTIVATION
\section{Motivation}
A 0/1 Knapsack problem is most often described as how does a theif steal the most valuable items with the amount of space he has available to carry them. However, the applications of a 0/1 Knapsack far beyond just this. When solved, a 0/1 Knapsack problem will render the optimal solution, whether it be how one should construct their schedule so they can get the most done in one day, or how to pack a truck such that the fewest amount of trips are taken.

%% END MOTIVATION

%% BEGIN BACKGROUND
\section{Background}
Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force. It is not pratical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers. However, there are other methods beside brute force that allow for the correct or 'good-enought' solution to be found. One of these methods is known as dynamic programming, which trades off speed for space. With dynamic programming the correct solution can be found in $ \theta{n*m} $ time, where n is the amount of items and m is the size of the knapsack. Another method for solving this type of problem is called a greedy algorithm, which works by making the best decision on which item to take locally. A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta{n} $ time. Unfortunately a greedy solution is not always correct and therefore highly unreliable.

%% END BACKGROUND

%% BEGIN PROCEDURE
\section{Procedure}
A multidimentional array insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
\newline
\textbf{Insertion Sort Pre-Condition}: A is an unsorted non-empty array of non-empty arrays containing a comparable data type with a natural order such that v is an index value of the inner array.
\newline
\textbf{Insertion Sort Post-Condition}: A' is a permutation of A that is in strictly non-increasing order.
\begin{algorithm}
\caption {\textsc{Insertion-Sort}(A, v)}
\label{algo:insertionsort}
\begin{algorithmic}[1]
\Procedure{Insertion-Sort}{A, v}
\If{$A.length < 2$}
\State{\Return{$A$}}
\EndIf
\State{$i = 2$}
\While{$i$ upto $A.length$}
	\State{$key = A[i][v]$}
	\State{$a = A[i]$}
	\State{$j = i - 2$}
	\While{$j$ downto $1$ and $key > A[j][v]$}
		\State{$A[j+1] = A[j]$}
		\State{$j = j - 1$}
	\EndWhile
	\State{$A[j+1] = a$}
	\State{$i = i + 1$}
\EndWhile
\State{\Return{$A$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 ... i - 1].
\newline 
\textbf{Insertion Sort Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
\newline
\textbf{Insertion Sort Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
\newline
\textbf{Insertion Sort Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated. The negation of the guard implies that A'[1 ... i - 1] contains all the elements in A[1 ... i - 1].
\newline
\newline
\textbf{Insertion Sort Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
\newline
\textbf{Insertion Sort Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
\newline
\textbf{Insertion Sort Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
\newline
\textbf{Insertion Sort Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ... j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant. 
\newline
\newline
\textbf{Insertion Sort Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order. This satisfies the post condition.
\newline
\newline

A greedy solution to a 0/1 knapsack problem can be implemented in a variety of languages using the pseudocode in Algorithm 2.
\newline
\textbf{Greedy Knapsack Pre-Condition}: $Weights$ and $Prices$ both have in them $n$ number of elements
\newline
\textbf{Greedy Knapsack Post-Condition}: The returned value will be a reasonable solution for the largest value of price combinations such that the aggregate of the corresponding weights does not exceed knapsack capacity $c$.
\begin{algorithm}
\caption {\textsc{Greedy-Knapsack}(n, Weights, Prices, c)}
\label{algo:greedy}
\begin{algorithmic}[1]
\Procedure{Greedy-Knapsack}{n, Weights, Prices, c}
\If{$n == 1$}
	\State{\Return{$Prices[1]$}}
\EndIf
\If{$n <= 0$}
	\State{\Return{$0$}}
\EndIf
\State{$profit = 0$}
\State{$ratio = new Array$}
\For{$v = 1$ upto $n$}
	\State{$a = Prices[v] / Weights[v]$}
	\State{$ratio[v] = a$}
\EndFor
\State{$ratio = Insertion-Sort(ratio, 1)$}
\State{$i = 1$}
\While{$c > 0$ and $i < n$}
	\If{$c - ratio[i][3] >= 0$}
		\State{$profit = profit + ratio[i][2]$}
		\State{$c = c - ratio[i][3]$}
	\EndIf
	\State{$i = i + 1$}
\EndWhile
\State{\Return{$profit$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
%% TODO: Write invariants for greedy algorithm here
\newline
\newline

A dynamic solution to 0/1 knapsack can be implemented in a variety of languages using the pseudocode provided in Algorithm 3.
\newline
\textbf{Dynamic Knapsack Pre-Condition}: $Weights$ and $Prices$ both have $n$ number of elements
\newline
\textbf{Dynamic Knapsack Post-Condition}: The returned value will be the correct solution for the largest value of price combinations such that the aggregate of the corresponding weights does not exceed knapsack capacity $c$.
\begin{algorithm}
\caption {\textsc{Dynamic-Knapsack}(n, Weights, Prices, c)}
\label{algo:dynamic}
\begin{algorithmic}[1]
\Procedure{Dynamic-Knapsack}{n, Weights, Prices, c}
\If{$n == 1$}
	\State{\Return{$Prices[1]$}}
\EndIf
\If{$n <= 0$}
	\State{\Return{$0$}}
\EndIf
\State{$tab[n][c] = new Nested Array$}
\For{$x = 1$ upto $n$}
	\For{$y = 1$ upto $c$}
		\State{$tab[x][y] = 0$}
	\EndFor
\EndFor
\For{$i = 1$ upto $n$}
	\For{$j = 0$ upto $c$}
		\If{$Weights[i] <= j$ and $Prices[i] + tab[i][j - Weights[i]] > tab[i][j]$}
			\State{$tab[i+1][j] = Price[i] + tab[i][j - weights[i]]$}
		\EndIf
		\If{$!$$(Weights[i] <= j)$ and $!$$(Prices[i] + tab[i][j - Weights[i]] > tab[i][j])$}
			\State{$tab[i+1][j] = tab[i][j]$}
		\EndIf
	\EndFor
\EndFor
\State{\Return{$tab[n][c]$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
%% TODO: Write pseudocode invariants here for Dynamic solution
\newline
\newline

%% END PROCEDURE

%% BEGIN TESTING
\section{Testing}
\subsection{Testing Plan and Results}
All arrays used in testing are Java ArrayList<Integer> except for $tab$ in the dynamic solution, which is a primitive Java nested array. All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgwick and Kevin Wayne \cite{stopwatch}
It is important to note that the stopwatch lass used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
When the number of elements was below 100, the size of the knapsack was 50; when values of 100 and greater were used the size of the knapsack was $2n$ where $n$ is the number of elements. For example, if the number of elements was 100, the size of the knapsack was 200, in order so the knapsack was sure to be able to make reasonable comparisons.
In the table below, times are given as averages out of 10 trials in milliseconds.
\newline
\captionof{table}{Greedy Solution Real-Time Run-Time}
\begin{center}
\begin{tabular}{|l|l|l|}
\hline Number of Elements & Size of Knapsack & Time \\
\hline 0 & 50 & 0.0 \\
\hline 1 & 50 & 0.0 \\
\hline 10 & 50 & 0.0006 \\
\hline 100 & 200 & 0.0043 \\
\hline 1000 & 2000 & 0.0231 \\
\hline 10000 & 20000 & 0.252 \\
\hline 100000 & 200000 & 59.1176 \\
\hline
\end{tabular}
\end{center}
\captionof{table}{Dynamic Solution Real-Time Run-Tim}
\begin{center}
\begin{tabular}{|l|l|l|}
\hline 0 & 50 & 0.0001 \\
\hline 1 & 50 & 0.0 \\
\hline 10 & 50 & 0.0002 \\
\hline 100 & 200 & 0.0119 \\
\hline 1000 & 2000 & 0.0356 \\
\hline 10000 & 20000 & 1.9243 \\
\hline
\end{tabular}
\end{center}
%% TODO: You need to implement a testing plan testing the correctness of the two algorithms!

\subsection{Problems Encountered}
One major issue encountered during the development of the dynamic programming solution was that when the second pair of loops would make calls to any of ArrayLists (including $tab$, which at the time was also an ArrayList implementation) there would be an out of bounds error because of a negative index value. The out of bounds error was eminating from the conditional in the second nested loop. The cause for this error is still unknown. To aid the issue $tab$ was changed from a nested ArrayList to a primitive nested Java array.

Another issue encountered during the development of this program was executing Java $assert$ statements. When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'. Removing the parenthesis wrapping the assertion fixed this issue.

When the dynamic solution was executed with values of anything above $n = 10000$ and $c = 20000$ Java gave a "Out Of Memory Error: Java heap space". This can be attributed the extreme number of elements being stored in $tab$. The computer's system monitor reported RAM usage at 95\%. Unfortunately, this issue cannot be aided.

The last issue encountered during the development process of this program was generic programming. The program would not compile because not all Objects have a compareTo method. Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue. \cite{erik}

%% END TESTING

%% BEGIN EXPERIMENTAL ANALYSIS
\section{Experimental Analysis}
%%TODO: write this
\begin{figure}[!]
\begin{center}
\includegraphics[scale=.55]{test-results.pdf}
\end{center}
\captionof{figure}{Java Implementation of Greedy and Dynamic Solutions to 0/1 Knapsack Run-Time Analysis}
\label{fig:runtime}
\end{figure}

%% END EXPERIMENTAL ANALYSIS

%% BEGIN CONCLUSION
\section{Conclusion}
%%TODO: write this

%% END CONCLUSION

\newpage

%% BEGIN REFERNCES
\section*{References}

%% \nocite{*}
%% \bibliographystyle{IEEEtran}
%% \bibliography{bib}

%% END REFERENCES

\newpage

%% BEGIN APPENDIX
\section*{Appendix}
\lstinputlisting[caption=Driver,
label={Driver.java},
breaklines=true,
]{Code/Driver.java}
\lstinputlisting[caption=Knapsack,
label={Knapsack.java},
breaklines=true,
]{Code/Knapsack.java}
\lstinputlisting[caption=Sort,
label={Sort.java},
breaklines=true,
]{Code/Sort.java}

The class Stopwatch has been altered from its original form.
\newline
\lstinputlisting[caption=Stopwatch,
label={Stopwatch.java},
breaklines=true,
]{Code/Stopwatch.java}
\lstinputlisting[caption=Debug,
label={Debug.java},
breaklines=true,
]{Code/Debug.java}
\lstinputlisting[caption=DummyData,
label={DummyData.java},
breaklines=true,
]{Code/DummyData.java}
%% END APPENDIX

%% END DOCUMENT
\end{document}