### 2015-03-13 Programming Project C/01knapsack.tex
should be complete

%% Based on a TeXnicCenter-Template by Tino Weinkauf.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% HEADER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[onecolumn, 11pt, conference, compsocconf]{IEEEtran}
\documentclass[onecolumn, 12pt, article]{IEEEtran}
\usepackage{times,color,amsmath,amssymb,amsthm,comment,graphicx,cite,multirow}
%\documentclass[letterpaper,oneside,11pt]{letter}
% Alternative Options:
%	Paper Size: a4paper / a5paper / b5paper / letterpaper / legalpaper / executivepaper
% Duplex: oneside / twoside
% Base Font Size: 10pt / 11pt / 12pt

%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[USenglish]{babel} %francais, polish, spanish, ...
\usepackage[T1]{fontenc}
\usepackage[ansinew]{inputenc}

\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}

%% Packages for Graphics & Figures %%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{graphicx} %%For loading graphic files
%\usepackage{subfig} %%Subfigures inside a figure
%\usepackage{tikz} %%Generate vector graphics from within LaTeX

%% Please note:
%% Images can be included using \includegraphics{filename}
%% resp. using the dialog in the Insert menu.
%% 
%% The mode "LaTeX => PDF" allows the following formats:
%%   .jpg  .png  .pdf  .mps
%% 
%% The modes "LaTeX => DVI", "LaTeX => PS" und "LaTeX => PS => PDF"
%% allow the following formats:
%%   .eps  .ps  .bmp  .pict  .pntg

%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{array,MnSymbol}

%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\singlespacing        %% 1-spacing (default)
%\onehalfspacing       %% 1,5-spacing
%\doublespacing        %% 2-spacing

%% Other Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{a4wide} %%Smaller margins = more text per page.
\usepackage{fancyhdr} %%Fancy headings
\usepackage{listings}
\usepackage{capt-of}

%
% Theorem like environments
%
\newtheorem{problem}{Problem}%
%\numberwithin{problem}{section}
\newtheorem{theorem}{Theorem}%
\newtheorem{acknowledgment}{Acknowledgment}%
%\newtheorem{algorithm}{Algorithm}%
\newtheorem{assumption}{Assumption}%
\newtheorem{axiom}{Axiom}%
\newtheorem{case}{Case}%
\numberwithin{case}{problem}
\newtheorem{claim}{Claim}%
\newtheorem{conclusion}{Conclusion}
\newtheorem{condition}{Condition}
\numberwithin{condition}{problem}
\numberwithin{condition}{subsection}
\newtheorem{conjecture}{Conjecture}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}
\newtheorem{definition}{Definition}
\numberwithin{definition}{section}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}%
\newtheorem{lemma}{Lemma}%
\newtheorem{notation}{Notation}%
\theoremstyle{remark}
\newtheorem{question}{Question}%
\numberwithin{question}{problem}
\theoremstyle{plain}
\newtheorem{answer}{Answer}%
\numberwithin{answer}{problem}
\newtheorem{proposition}{Proposition}%
\newtheorem{remark}{Remark}%
\newtheorem{solution}{Solution}%
\numberwithin{solution}{section}
\newtheorem{summary}{Summary}%
\numberwithin{equation}{section}%
\newtheorem{option}{Option}

\raggedbottom
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Remarks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Note:
% 1. Edit the used packages and their options (see above).
% 2. If you want, add a BibTeX-File to the project
%    (e.g., 'literature.bib').
% 3. Happy TeXing!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Options / Modifications
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{options} %You need a file 'options.tex' for this
%% ==> TeXnicCenter supplies some possible option files
%% ==> with its-templates (File | New from Template...).

%% BEGIN DOCUMENT
\begin{document}

%% Title Page
\title{0/1 Knapsack}
\author{Preston Stosur-Bassett}
\date{March 11, 2015}
\maketitle

\pagestyle{fancy}
\fancyhead[R]{0/1 Knapsack in Java, page \thepage}
\fancyhead[L]{Preston Stosur-Bassett}

%% BEGIN ABSTRACT
\begin{abstract}
The present study investigates two implementations for a 0/1 Knapsack problem, a greedy solution, and a dynamic solution. The dynamic solution will always result in the correct answer but takes $\theta{n*m}$ time to complete. The greedy solution will generally result in a reasonable answer, and even sometimes the correct answer, and take $\theta{n}$ time to complete. Which solution to use will depend on what is most important to the problem at hand, speed, or accuracy.

%% END ABSTRACT
\end{abstract}

%% BEGIN MOTIVATION
\section{Motivation}
A 0/1 Knapsack problem is most often described as how does a thief steal the most valuable items with the amount of space he has available to carry them. However, the applications of a 0/1 Knapsack far beyond just this. When solved, a 0/1 Knapsack problem will render the optimal solution, whether it be how one should construct their schedule so they can get the most done in one day, or how to pack a truck such that the fewest amount of trips are taken.

%% END MOTIVATION

%% BEGIN BACKGROUND
\section{Background}
Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force. It is not practical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers. However, there are other methods beside brute force that allow for the correct or 'good-enough' solution to be found. One of these methods is known as dynamic programming, which trades off speed for space. With dynamic programming the correct solution can be found in $ \theta{n*m} $ time, where n is the amount of items and m is the size of the knapsack. Another method for solving this type of problem is called a greedy algorithm, which works by making the best decision on which item to take locally. A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta{n} $ time. Unfortunately a greedy solution is not always correct and therefore highly unreliable.

%% END BACKGROUND

%% BEGIN PROCEDURE
\section{Procedure}
A multidimensional array insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
\newline
\textbf{Insertion Sort Pre-Condition}: A is an unsorted non-empty array of non-empty arrays containing a comparable data type with a natural order such that v is an index value of the inner array.
\newline
\textbf{Insertion Sort Post-Condition}: A' is a permutation of A that is in strictly non-increasing order.
\begin{algorithm}
\caption {\textsc{Insertion-Sort}(A, v)}
\label{algo:insertionsort}
\begin{algorithmic}[1]
\Procedure{Insertion-Sort}{A, v}
\If{$A.length < 2$}
\State{\Return{$A$}}
\EndIf
\State{$i = 2$}
\While{$i$ upto $A.length$}
	\State{$key = A[i][v]$}
	\State{$a = A[i]$}
	\State{$j = i - 2$}
	\While{$j$ downto $1$ and $key > A[j][v]$}
		\State{$A[j+1] = A[j]$}
		\State{$j = j - 1$}
	\EndWhile
	\State{$A[j+1] = a$}
	\State{$i = i + 1$}
\EndWhile
\State{\Return{$A$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 ... i - 1].
\newline 
\textbf{Insertion Sort Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
\newline
\textbf{Insertion Sort Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
\newline
\textbf{Insertion Sort Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated. The negation of the guard implies that A'[1 ... i - 1] contains all the elements in A[1 ... i - 1].
\newline
\newline
\textbf{Insertion Sort Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
\newline
\textbf{Insertion Sort Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
\newline
\textbf{Insertion Sort Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
\newline
\textbf{Insertion Sort Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ... j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant. 
\newline
\newline
\textbf{Insertion Sort Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order. This satisfies the post condition.
\newline
\newline

A greedy solution to a 0/1 knapsack problem can be implemented in a variety of languages using the pseudocode in Algorithm 2.
\newline
\textbf{Greedy Knapsack Pre-Condition}: $Weights$ and $Prices$ both have in them $n$ number of elements
\newline
\textbf{Greedy Knapsack Post-Condition}: The returned value will be a reasonable solution for the largest value of price combinations such that the aggregate of the corresponding weights does not exceed knapsack capacity $c$.
\begin{algorithm}
\caption {\textsc{Greedy-Knapsack}(n, Weights, Prices, c)}
\label{algo:greedy}
\begin{algorithmic}[1]
\Procedure{Greedy-Knapsack}{n, Weights, Prices, c}
\If{$n == 1$ and $Weights[1] <= c$}
	\State{\Return{$Prices[1]$}}
\EndIf
\If{$n <= 0$}
	\State{\Return{$0$}}
\EndIf
\State{$profit = 0$}
\State{$ratio = new Array$}
\For{$v = 1$ upto $n$}
	\State{$a = Prices[v] / Weights[v]$}
	\State{$ratio[v] = a$}
\EndFor
\State{$ratio = Insertion-Sort(ratio, 1)$}
\State{$i = 1$}
\While{$c > 0$ and $i < n$}
	\If{$c - ratio[i][3] >= 0$}
		\State{$profit = profit + ratio[i][2]$}
		\State{$c = c - ratio[i][3]$}
	\EndIf
	\State{$i = i + 1$}
\EndWhile
\State{\Return{$profit$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Greedy Solution First-Loop Invariant}: ratio[1 ... v-1] has the same number of non-null elements in it as both Prices[1 ... v-1] and Weights[1 ... v-1].
\newline
\textbf{Greedy Solution First-Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because $v=1$, and ratio[1 ... v-1], Prices[1 ... v-1] and Weights[0 ... v-1] all do not exist, and is therefore vacuously true.
\newline
\textbf{Greedy Solution First-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because each time the loop runs, v is incremented by 1, and exactly 1 element is added to ratio, therefore ratio[0 ... v-1] will have the exact same number of elements as both Prices[1 ... v-1] and Weights[1 ... v-1].
\newline
\textbf{Greedy Solution First-Loop Invariant Termination}: After the termination of the loop the invariant holds true because v has been incremented by 1 during each iteration of the loop, and ratio has gained 1 element during each iteration of the loop, therefore ratio[1 ... v-1] contains the exact same number of elements as both Prices[1 ... v-1] and Weights[1 ... v-1].
\newline
\newline
\textbf{Greedy Solution Second-Loop Invariant}: The aggregate of the weights corresponding to the prices included in $profit$ are less than or equal to $c$.
\newline
\textbf{Greedy Solution Second-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because no prices have been included in $profit$.
\newline
\textbf{Greedy Solution Second-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because only items whose weight is less than or equal to the size of $c$ are added to $profit$.
\newline
\textbf{Greedy Solution Second-Loop Invariant Termination}: After the termination of the loop, the invariant holds true because only items whose weight was less than or equal to the value of $c$ have been added to $profit$.
\newline
\newline

A dynamic solution to 0/1 knapsack can be implemented in a variety of languages using the pseudocode provided in Algorithm 3.
\newline
\textbf{Dynamic Knapsack Pre-Condition}: $Weights$ and $Prices$ both have $n$ number of elements
\newline
\textbf{Dynamic Knapsack Post-Condition}: The returned value will be the correct solution for the largest value of price combinations such that the aggregate of the corresponding weights does not exceed knapsack capacity $c$.
\begin{algorithm}
\caption {\textsc{Dynamic-Knapsack}(n, Weights, Prices, c)}
\label{algo:dynamic}
\begin{algorithmic}[1]
\Procedure{Dynamic-Knapsack}{n, Weights, Prices, c}
\If{$n == 1$ and $Weights[1] <= c$}
	\State{\Return{$Prices[1]$}}
\EndIf
\If{$n <= 0$}
	\State{\Return{$0$}}
\EndIf
\State{$tab[n][c] = new Nested Array$}
\For{$x = 1$ upto $n$}
	\For{$y = 1$ upto $c$}
		\State{$tab[x][y] = 0$}
	\EndFor
\EndFor
\For{$i = 1$ upto $n$}
	\For{$j = 0$ upto $c$}
		\If{$Weights[i] <= j$ and $Prices[i] + tab[i][j - Weights[i]] > tab[i][j]$}
			\State{$tab[i+1][j] = Price[i] + tab[i][j - weights[i]]$}
		\EndIf
		\If{$!$$(Weights[i] <= j)$ and $!$$(Prices[i] + tab[i][j - Weights[i]] > tab[i][j])$}
			\State{$tab[i+1][j] = tab[i][j]$}
		\EndIf
	\EndFor
\EndFor
\State{\Return{$tab[n][c]$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Dynamic Solution First Outer-Loop Invariant}: tab[1 .. x-1] has the same number of non-null elements in it as both Weights[1 ... x - 1] and Prices[1 ... x-1].
\newline
\textbf{Dynamic Solution First Outer-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because tab[0 ... x-1], weights[0 ... x-1] and prices[0 ... x-1] all do not exist.
\newline
\textbf{Dynamic Solution First Outer-Loop Invariant Maintenance}: At the beginning of each iteration because x is incremented by one at the same rate that one element is added to tab.
\newline
\textbf{Dynamic Solution First Outer-Loop Invariant Termination}: At the termination of the loop the invariant holds true because x has been incremented at the same rate that an element has been added to tab.
\newline
\newline
\textbf{Dynamic Solution First Inner-Loop Invariant}: tab[x] has y number of non-null elements in it.
\newline
\textbf{Dynamic Solution First Inner-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true because y = 0 and tab[x] has 0 elements in it.
\newline
\textbf{Dynamic Solution First Inner-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because because y is incremented at the same rate that elements are added to tab[x].
\newline
\textbf{Dynamic Solution First Inner-Loop Invariant Termination}: After the termination of the loop the invariant holds true because elements have been added to tab[x] at the same rate in which y has been incremented.
\newline
\newline
\textbf{Dynamic Solution Second Outer-Loop Invariant}: tab[i][j] is larger than or equal to tab[i-1][j]
\newline
\textbf{Dynamic Solution Second Outer-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because tab[-1] does not exist.
\newline
\textbf{Dynamic Solution Second Outer-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because during the loop the values are incremented, but they are never decreased.
\newline
\textbf{Dynamic Solution Second Outer-Loop Invariant Termination}: After the loop has terminated the invariant holds true because the values in tab have never been decreased, only increased, and therefore tab[i][j] will be greater than or equal to tab[i-1][j]
\newline
\newline
\textbf{Dynamic Solution Second Inner-Loop Invariant}: tab[i][j] is always equal to or greater than tab[i][j-1]
\newline
\textbf{Dynamic Solution Second Inner-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because tab[i][j-1] does not exist.
\newline
\textbf{Dynamic Solution Second Inner-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because the values in tab have never been decreased, only increased, therefore tab[i][j] is going to be larger than or equal to tab[i][j-1].
\newline
\textbf{Dynamic Solution Second Inner-Loop Invariant Termination}: After the termination of the loop the invariant holds true because the values in tab have never been decremented, only incremented.
\newline
\newline

%% END PROCEDURE

%% BEGIN TESTING
\section{Testing}
\subsection{Testing Plan and Results}
All arrays used in testing are Java ArrayList<Integer> except for $tab$ in the dynamic solution, which is a primitive Java nested array. All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgwick and Kevin Wayne \cite{stopwatch}
It is important to note that the stopwatch class used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
The size of elements used in testing are as follows: For $n < 100$, $c = 50$. For $n >= 100$, $c = 2n$ where $n$ represents the number of items, and $c$ represents the capacity of the knapsack.
\newline
\captionof{table}{Greedy Solution Real-Time Run-Time}
\begin{center}
\begin{tabular}{|l|l|l|}
\hline Number of Elements & Size of Knapsack & Time \\
\hline 0 & 50 & 0.0 \\
\hline 1 & 50 & 0.0 \\
\hline 10 & 50 & 0.0006 \\
\hline 100 & 200 & 0.0043 \\
\hline 1000 & 2000 & 0.0231 \\
\hline 10000 & 20000 & 0.252 \\
\hline 100000 & 200000 & 59.1176 \\
\hline
\end{tabular}
\end{center}
\captionof{table}{Dynamic Solution Real-Time Run-Time}
\begin{center}
\begin{tabular}{|l|l|l|}
\hline Number of Elements & Size of Knapsack & Time \\
\hline 0 & 50 & 0.0001 \\
\hline 1 & 50 & 0.0 \\
\hline 10 & 50 & 0.0002 \\
\hline 100 & 200 & 0.0119 \\
\hline 1000 & 2000 & 0.0356 \\
\hline 10000 & 20000 & 1.9243 \\
\hline
\end{tabular}
\end{center}
The percent error of greedy solutions is shown in Table 3. Percent error was calculated by the percent error formula: $\frac{ | Greedy Solution - Dynamic Solution | }{ | Dynamic Solution | } * 100$. The percentages shown in Table 3 are average percent errors taken from trials of 10.
\captionof{table}{Dynamic vs Greedy Solutions Accuracy Comparison}
\begin{center}
\begin{tabular}{|l|l|l|l}
\hline Number of Elements & Size of Knapsack & Percent Error \\
\hline 0 & 50 & 0\% \\
\hline 1 & 50 & 0\% \\
\hline 10 & 50 & 5.3\% \\
\hline 100 & 200 & 0.9\% \\
\hline 1000 & 2000 & 0.032887\% \\
\hline 10000 & 20000 & 0.00555725\% \\
\hline
\end{tabular}
\end{center}

\subsection{Problems Encountered}
One major issue encountered during the development of the dynamic programming solution was that when the second pair of loops would make calls to any of ArrayLists (including $tab$, which at the time was also an ArrayList implementation) there would be an out of bounds error because of a negative index value. The out of bounds error was emanating from the conditional in the second nested loop. The cause for this error is still unknown. To aid the issue $tab$ was changed from a nested ArrayList to a primitive nested Java array.

Another issue encountered during the development of this program was executing Java $assert$ statements. When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'. Removing the parenthesis wrapping the assertion fixed this issue.

When the dynamic solution was executed with values of anything above $n = 10000$ and $c = 20000$ Java gave a "Out Of Memory Error: Java heap space". This can be attributed the extreme number of elements being stored in $tab$. The computer's system monitor reported RAM usage at 95\%. Unfortunately, this issue cannot be aided.

The last issue encountered during the development process of this program was generic programming. The program would not compile because not all Objects have a compareTo method. Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue. \cite{erik}

%% END TESTING

%% BEGIN EXPERIMENTAL ANALYSIS
\section{Experimental Analysis}
\begin{figure}[!]
\begin{center}
\includegraphics[scale=.55]{test-results.pdf}
\end{center}
\captionof{figure}{Java Implementation of Greedy and Dynamic Solutions to 0/1 Knapsack Run-Time Analysis}
\label{fig:runtime}
\end{figure}
The insertion sort, greedy solution and dynamic solution demonstrated in Algorithm 1, 2 and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 Ivybridge processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit with 8GB of DDR3 RAM.
The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta{n^2} $ where $ \theta{} $ represents the asymptotically tightly bound running time. \cite{textbook}

The greedy solution also follows a similar $ \theta{n^2} $ run time, see Figure 1. This is because an insertion sort was chosen to implement as the sorting algorithm used in the greedy solution. An insertion sort was chosen for the sorting algorithm because, while it is not the fastest, requires less axillary space. While run time is an important factor, the space required to run the algorithm is also very important, and should not be neglected. As seen in Table 2, the dynamic solution unavoidably required too much axillary space and Java gave an error when values of $n$ became increasingly large. It also important to note that the data graphed in Figure 1 was executed while the machine executed superfluous commands such as checking for system updates. To aid this issue, averages of 10 trails were taken and used in Figure 1.

The run time for the dynamic solution as seen in figure one $ \theta{n*c} $ where n is the number of elements and c is the size of the knapsack. In this implementation for values of $n$ greater than or equal to 100, a value of $2n$ was used for $c$. That would make the run time of the specific implementation of the dynamic solution $2n^2$ but after dropping constants becomes $ \theta{n^2} $. In Figure 1, the reason that the dynamic solution takes so much more time to complete is because there is a constant of 2 for $n^2$ which can be attributed to doubling the real time it took to return the results. Because the dynamic solution works by storing answers originally found in axillary space, the values of $n$ cannot be very high without causing a Java error, as seen in Table 2.

In Table 3 there looks to be a negative correlation between the size of $n$ and the percent error. This negative correlation can be attributed to the fact that the percent error formula takes the difference between the two results and divides it by the correct result. For example, if the result for a greedy solution was 2 and the result for the dynamic solution was 3 the formula would read $ \frac{2}{3} * 100 $ rendering a 66.66\% error, however the difference between the two is only 2. If the values were 101 and 103 the percent error formula would read $ \frac{2}{103} * 100 $ and renders 1.9\% error. The second value is substantially lower, even though the difference between the two numbers remains the same.

%% END EXPERIMENTAL ANALYSIS

%% BEGIN CONCLUSION
\section{Conclusion}
Although in this specific implementation, both the dynamic solution and the greedy solution had a run-time of $\theta{n^2}$, it can be possible for the dynamic solution to have a run-time of $\theta{n*c}$ and the greedy solution to have a run-time of $\theta{n}$. The best algorithm to use for a given problem is really a subjective decision and depends on two things, 1. how accurate does the answer to the problem really need to be, and 2. what is the time requirement to find out what the answer is.

%% END CONCLUSION

\newpage

%% BEGIN REFERNCES
%%\section*{References}

\nocite{*}
\bibliographystyle{IEEEtran}
\bibliography{bib}

%% END REFERENCES

\newpage

%% BEGIN APPENDIX
\section*{Appendix}
\lstinputlisting[caption=Driver,
label={Driver.java},
breaklines=true,
]{Code/Driver.java}
\lstinputlisting[caption=Knapsack,
label={Knapsack.java},
breaklines=true,
]{Code/Knapsack.java}
\lstinputlisting[caption=Sort,
label={Sort.java},
breaklines=true,
]{Code/Sort.java}

The class Stopwatch has been altered from its original form.
\newline
\lstinputlisting[caption=Stopwatch,
label={Stopwatch.java},
breaklines=true,
]{Code/Stopwatch.java}
\lstinputlisting[caption=Debug,
label={Debug.java},
breaklines=true,
]{Code/Debug.java}
\lstinputlisting[caption=DummyData,
label={DummyData.java},
breaklines=true,
]{Code/DummyData.java}
%% END APPENDIX

%% END DOCUMENT
\end{document}